The evolution of gene expression in the term placenta of viviparous mammals

### 2016-04-18 manuscript/manuscript.Rnw
fix the commas and input.read.length issue

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","placenta_go_results_mf","hugo_nomenclature","gene_annotation")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf",
                              "hugo_nomenclature",
                              "gene_annotation"
                              ))) {
    load(a)
}
combined.fpkm[,species:=factor(species,levels=names(species.names))]
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
star.logs.wide[,`% bad reads`:=(as.numeric(`% of reads mapped to too many loci`)+
                            as.numeric(`% of reads unmapped: other`)+
                            as.numeric(`% of reads unmapped: too many mismatches`)+
                            as.numeric(`% of reads unmapped: too short`))]
star.logs.wide[,`% good reads`:=100-`% bad reads`]
star.logs.wide[,input.reads:=as.numeric(input.reads)]
setkey(hugo.nomenclature,"Approved Symbol")
@

@@

Supplemental \cref{fig:transcriptome_statistics} shows the RNAseq read and
alignment statistics for all species. For transcriptomes sequenced
with 75 bp %
%
(\Sexpr{species(star.logs.wide[input.read.length > 70 & input.read.length < 80,unique(species)])})
%
,
the number of trimmed reads ranged from %
\Sexpr{star.logs.wide[input.read.length > 70 & input.read.length < 80,min(as.numeric(input.reads))]} 
%
to %
\Sexpr{star.logs.wide[input.read.length > 70 & input.read.length < 80,max(as.numeric(input.reads))]}
with an average of \Sexpr{star.logs.wide[input.read.length > 70 & input.read.length < 80,mean(`% good reads`)]}\% of reads 
mapping to the respective reference genome (\cref{tab:ref_genome_aligned_species}). 
%
For transcriptomes sequenced with 150 bp, %
% The mismatch here is because we're reporting trimmed reads here, not
% full length reads
(\Sexpr{species(star.logs.wide[input.read.length > 120 & input.read.length < 160,unique(species)])})
%
,
the number of trimmed reads ranged from %
\Sexpr{star.logs.wide[input.read.length > 120 & input.read.length < 160,min(as.numeric(input.reads))]} 
%
to %
\Sexpr{star.logs.wide[input.read.length  > 120 & input.read.length < 160,max(as.numeric(input.reads))]}
%
with an average of \Sexpr{star.logs.wide[input.read.length > 120 & input.read.length < 160,mean(`% good reads`)]}\% %
of reads mapping to the
respective reference genome (\cref{tab:ref_genome_aligned_species}).
%
%
For transcriptomes sequenced with 35 bp, %
% The mismatch here is because we're reporting trimmed reads here, not
% full length reads
(\Sexpr{species(star.logs.wide[input.read.length < 40 ,unique(species)])})
%
,
the number of trimmed reads ranged from %
\Sexpr{star.logs.wide[input.read.length < 40 ,min(as.numeric(input.reads))]} 
%
to %
\Sexpr{star.logs.wide[input.read.length  < 40  ,max(as.numeric(input.reads))]}
%
with an average of \Sexpr{star.logs.wide[input.read.length < 40, mean(`% good reads`)]}\%
of reads mapping to the
respective reference genome (\cref{tab:ref_genome_aligned_species}).
The top 50 expressed genes in each genome analyzed are shown in
\cref{tab:top50_human_placenta,tab:top50_bonobo_placenta,tab:top50_spider_monkey_placenta,tab:top50_dog_placenta,tab:top50_cow_placenta,tab:top50_elephant_placenta,tab:top50_opossum_placenta,tab:top50_mouse_placenta,tab:top50_n_galili_placenta,tab:top50_s_carmeli_placenta,tab:top50_pig_placenta,tab:top50_horse_placenta,tab:top50_sheep_placenta}
and all genes in Supplemental \cref{tab:all_ateles_fusciceps_placenta,tab:all_homo_sapiens_placenta,tab:all_pan_paniscus_placenta,tab:all_mus_musculus_placenta,tab:all_spalax_galili_placenta,tab:all_spalax_carmeli_placenta,tab:all_bos_taurus_placenta,tab:all_ovis_aries_placenta,tab:all_sus_scrofa_placenta,tab:all_equus_caballus_placenta,tab:all_canis_familiaris_placenta,tab:all_loxodonta_africana_placenta,tab:all_dasypus_novemcinctus_placenta,tab:all_monodelphis_domestica_placenta}.

>>>>

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","placenta_go_results_mf","hugo_nomenclature","gene_annotation")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf",
                              "hugo_nomenclature",
                              "gene_annotation"
                              ))) {
    load(a)
}
combined.fpkm[,species:=factor(species,levels=names(species.names))]
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
star.logs.wide[,`% bad reads`:=(as.numeric(`% of reads mapped to too many loci`)+
                            as.numeric(`% of reads unmapped: other`)+
                            as.numeric(`% of reads unmapped: too many mismatches`)+
                            as.numeric(`% of reads unmapped: too short`))]
star.logs.wide[,`% good reads`:=100-`% bad reads`]
star.logs.wide[,input.reads:=as.numeric(input.reads)]
star.logs.wide[,input.read.length:=as.numeric(input.read.length)]
setkey(hugo.nomenclature,"Approved Symbol")
@

@@

Supplemental \cref{fig:transcriptome_statistics} shows the RNAseq read and
alignment statistics for all species. For transcriptomes sequenced
with 75 bp %
%
(\Sexpr{species(star.logs.wide[input.read.length > 70 & input.read.length < 80,unique(species)])}),%
%
the number of trimmed reads ranged from %
\Sexpr{star.logs.wide[input.read.length > 70 & input.read.length < 80,min(as.numeric(input.reads))]} 
%
to %
\Sexpr{star.logs.wide[input.read.length > 70 & input.read.length < 80,max(as.numeric(input.reads))]}
with an average of \Sexpr{star.logs.wide[input.read.length > 70 & input.read.length < 80,mean(`% good reads`)]}\% of reads 
mapping to the respective reference genome (\cref{tab:ref_genome_aligned_species}). 
%
For transcriptomes sequenced with 150 bp %
% The mismatch here is because we're reporting trimmed reads here, not
% full length reads
(\Sexpr{species(star.logs.wide[input.read.length > 120 & input.read.length < 160,unique(species)])}),
%
the number of trimmed reads ranged from %
\Sexpr{star.logs.wide[input.read.length > 120 & input.read.length < 160,min(as.numeric(input.reads))]} 
%
to %
\Sexpr{star.logs.wide[input.read.length  > 120 & input.read.length < 160,max(as.numeric(input.reads))]}
%
with an average of \Sexpr{star.logs.wide[input.read.length > 120 & input.read.length < 160,mean(`% good reads`)]}\% %
of reads mapping to the
respective reference genome (\cref{tab:ref_genome_aligned_species}).
%
%
For transcriptomes sequenced with 35 bp %
% The mismatch here is because we're reporting trimmed reads here, not
% full length reads
(\Sexpr{species(star.logs.wide[input.read.length < 40 ,unique(species)])}),
%
the number of trimmed reads ranged from %
\Sexpr{star.logs.wide[input.read.length < 40 ,min(as.numeric(input.reads))]} 
%
to %
\Sexpr{star.logs.wide[input.read.length  < 40  ,max(as.numeric(input.reads))]}
%
with an average of \Sexpr{star.logs.wide[input.read.length < 40, mean(`% good reads`)]}\%
of reads mapping to the
respective reference genome (\cref{tab:ref_genome_aligned_species}).
The top 50 expressed genes in each genome analyzed are shown in
\cref{tab:top50_human_placenta,tab:top50_bonobo_placenta,tab:top50_spider_monkey_placenta,tab:top50_dog_placenta,tab:top50_cow_placenta,tab:top50_elephant_placenta,tab:top50_opossum_placenta,tab:top50_mouse_placenta,tab:top50_n_galili_placenta,tab:top50_s_carmeli_placenta,tab:top50_pig_placenta,tab:top50_horse_placenta,tab:top50_sheep_placenta}
and all genes in Supplemental \cref{tab:all_ateles_fusciceps_placenta,tab:all_homo_sapiens_placenta,tab:all_pan_paniscus_placenta,tab:all_mus_musculus_placenta,tab:all_spalax_galili_placenta,tab:all_spalax_carmeli_placenta,tab:all_bos_taurus_placenta,tab:all_ovis_aries_placenta,tab:all_sus_scrofa_placenta,tab:all_equus_caballus_placenta,tab:all_canis_familiaris_placenta,tab:all_loxodonta_africana_placenta,tab:all_dasypus_novemcinctus_placenta,tab:all_monodelphis_domestica_placenta}.
